---
title: "Task 4"
output: html_notebook
---
join 3 data sets
2016 2015 timestamp data (unique ID?)2017 unique ID
```{r}
library(tidyverse)

candyclean <- read_csv("clean_data/clean_candy_data.csv")
```


What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)

```{r}
candyclean %>% 
  select(7:114) %>% # select columns containing rating for all candies  
  tally(!is.na(.), name = 'total_number_of_candy_ratings' )
```



j
What was the average age of people who are going out trick or treating and the average age of people 3. not going trick or treating?
age, trick or treat value

```{r}
candyclean %>% 
  select(2, 4) %>% #select colums needed
  filter(age > 0 & age < 120) %>% #filter ages used to sensible range
  group_by(going_out) %>% # group by going out trick or treating or not
  summarise_if(is.numeric, mean, na.rm=TRUE) #perform mean for ages on each category
```


For each of joy, despair and meh, which candy bar revived the most of these ratings?
aggregate 
```{r}
candyclean %>% 
  
```

How many people rated Starburst as despair?
```{r}
candyclean %>% 
  select(starburst) %>% 
  filter(str_detect(starburst, "DESPAIR")) %>% 
  tally(name = 'How many respondents despair over starburst')
  
```


For the next three questions, count despair as -1, joy as +1 and meh as 0.


What was the most popular candy bar by this rating system for each gender in the dataset?

```{r}
candy_rate <- candyclean %>%
  mutate(across(c(7:114), 
             ~ case_when(. == "DESPAIR" ~ -1,
                         . == "JOY" ~ 1,
                         . == "MEH" ~ 0
                         ))) %>% 
  select(3, 7:114) %>% 
  group_by(gender) %>%
  summarize_if(is.numeric, sum, na.rm=TRUE)
  
```

```{r}
candy_gender <- candy_rate %>% 
  pivot_longer(!gender, names_to = "candy_type", values_to = 'rating_total') %>% 
  arrange(desc(rating_total), group_by = gender) 


```



What was the most popular candy bar in each year?
```{r}
candy_gender %>% 
  group_by(gender, candy_type) %>%
  summarise(
    MaxMassByGender = max(rating_total, na.rm = T)
  ) %>%
  arrange(gender)

```


What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?
no  location data for 2015
join 2016 and 2017
```{r}

```


